---
title: "Lab 4_Report"
author: "Sarah Wicks"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Sarah Wicks

**BIO 781 - R Stats & Modeling**

2026-09-21

**Lab 4 - For Loops, Functions, and Bootstrapping in R**

```{r Code block for Lab 4, include=TRUE, results='hide'}

#create a set of 25 values in the normal distribution
x <- rnorm(25)
hist(x, freq = FALSE, ylim = c(0,2), xlim = c(-4,4), main = "25 observations")

#calculate the mean of that distribution 1000 times
mean.x <- replicate(1000, mean(rnorm(25)))
hist(mean.x, add = TRUE, col = "blue", freq = F)

#calculate the 95% interval of the data
sd(x)
(mean(x))+(1.96*(sd(x)))
(mean(x))-(1.96*(sd(x)))


#Calculate the 95% confidence interval on the mean
(mean(x))+(1.96*(sd(x)/sqrt(25)))
(mean(x))-(1.96*(sd(x)/sqrt(25)))

```

**Checkpoint 1: Using the equations above, what is the region where 95%
of the data points lie, and what is the region where there is a 95%
chance that the mean of the population exists within it? How do these
regions differ**

Standard deviation of x when n=25 is 1.1. The mean of x = -0.091. Using
the calculations given prior:

95% of the data points lie between (-2.254, 2.072)

The region with a 95% chance of containing the mean: (-0.524, 0.341)

These regions differ in that the first uses the standard deviation of
the whole sample and shows where most of the data should lie if
re-sampled, whereas the 95% CI on the mean uses the standard error to
calculate the range in which we can expect to find the mean of the
population.

**Checkpoint 2: How do these regions compare to the example in which we
drew 1000 samples from the distribution?**

95% Intervals of the data: (-1.928, 1.830)

95% CI on the mean: (-0.108, 0.0108)

The 95% intervals of the data are relatively similar, the 1000
observation example's window is slightly smaller. The 95% CI on the mean
of the sample with 1000 observations is much smaller, reflecting the
larger sample size's impact on the "trustworthiness" of the sample (as
indicated by the lower standard error in this sample).

**Checkpoint 3: How do we interpret this statistical test given that our
residuals are non-normal (and so the test isn’t valid)?**

Because we know that the residuals are not normally distributed, the
p-values and confidence intervals are not trustworthy to determine
significance. Different tests would need to be used to determine the
relationship's significance.

**Checkpoint 4: What does the bootstrap analysis suggest about our data?
What if we had done more bootstraps? Why is looking at the standard
error useless in a bootstrap?**

The bootstrap analysis suggests that the true nature of the slope
(relationship between growth rate and temp) is normally distributed, but
the sample size is too small to determine the distribution without
re-sampling. If we had done more bootstraps, it might show an even
stronger adherence to a normal distribution. Standard error in a
bootstrap isn't useful because it assumes a normal distribution, which
the underlying data for this bootstrap don't have.
